\LARGE \bf Informed sampling by Markov Chain Monte Carlo methods

### 2017-01-30 paper.tex
related work

Initially, 
when no path has yet to be found, 
the samples need to be drawn from the entire state space~$\calX$.
However, once a path $\gamma$ is produced,  algorithms that seek \emph{high-quality paths} can limit their sampling domain to a subset of~$\calX$  containing states that may produce higher-quality paths than~$\gamma$.
Following Gammell et al.~\cite{GSB14}, we call this subset the \emph{informed subset} and denote it $\Cinf$.

@@

Unfortunately, both rejection sampling and HRS may produce a large number of rejected samples especially in high-dimensional spaces.
This may cause the search algorithm to spend most of its time trying to produces samples in~$\Cinf$ rather than explore it.
In this paper, we suggest an alternative approach to sampling in the informed set \Cinf for a wide range of settings.
Our main insight is 
\os{this is a very important sentence!!!}
Specifically, our approach consists of two stages:
in the first, a random sample $x \in \calX$ is retracted to the boundary of~$\Cinf$ by running a gradient-descent algorithm;
in the second stage, this retracted sample $\tilde{x}$ is used to seed a Monte Carlo sampling chain which allows us to approximately produce samples uniformly \emph{within}~$\Cinf$.

Our approach requires that the system has a solution to the two-point boundary value problem (2pBVP)~\cite{} and that a gradient can be defined over the cost function.
Indeed, we demonstrate the efficiency of our approach on a wide variety of systems and show that it enables reducing the planning time by several orders of magnitute when compared to algorithms using rejection sampling or HRS.

@@

\subsection{Planning}
\label{subsec:planning}
Initial sampling-based algorithms such as RRT~\cite{LK01} and PRM~\cite{KSLO96} did not take into account the \emph{quality} of a path, given some optimization criteria, and only guaranteed to asymptotically return \emph{a} solution, if one exists.
Karaman and Frazzoli~\cite{KF11}, presented variants of PRM and RRT, named PRM* and RRT*, respectively that were shown to produce paths who's cost converges asymptotically to the minimal-cost path.
Additional algorithms followed, increasing the converges rate by various techniques such as lazy dynamic programming~\cite{GSB15, JSCP15}, relaxing optimality to near-optimality~\cite{DB14, SH16} and more.

Many of the algorithms mentioned require solving a two-point boundary value problem (2pBVP) to perform exact connections between vertices in the roadmap.

Xie et al.~\cite{XBPA15} use a variant of sequential quadratic programming (SQP) to solve 2pBVP and integrate it with BIT*~\cite{GSB15}.

Webb and van den Berg~\cite{WB13} use a fixed-final-state-free-final-time controller to solve the 2pBVP  with respect to a cost function that is allows for balancing between the duration of the trajectory and the expended control effort.

RGGS~\cite{SSH16}

Kinodynamic works:\\~\cite{KF10}

Perez et al.~\cite{PPKKL12} propose a variant of RRT* that automatically defines a distance metric and node extension method by locally linearizing
the domain dynamics and applying linear quadratic regulation (LQR).

Goretkin et al.~\cite{GPPK13}

\subsection{Sampling}
\label{subsec:sampling}
Hierarchical sampling~\cite{KTC16}.

Informed sampling~\cite{GSB14}

OBPRM~\cite{ABDJV98, YTEA12}

Medial axis~\cite{LTA03, YDLTA14}

\os{can our method be applied to produce samples on the medial axis?}

S. A. Wilmarth, N. M. Amato, and P. F. Stiller. MAPRM: A prob-
abilistic roadmap planner with sampling on the medial axis of the
free space. In Proc. IEEE Int. Conf. Robot. Autom. (ICRA), pages
1024–1031, 1999.

S. A. Wilmarth, N. M. Amato, and P. F. Stiller. Motion planning for
a rigid body using random networks on the medial axis of the free
space. In Proc. ACM Symp. on Computational Geometry (SoCG),
pages 173–180, 1999

heuristic~\cite{US03, SWT09}

\subsection{MCMC}
\label{subsec:mcmc}
Markov Chain Monte Carlo(MCMC)~\cite{ADDJ03} is a framework of sampling from a given probability distribution by constructing a Markov chain of random transition toward a desired distribution $ \pi (x) $ .
A set of samples are generated to approximate the desired distribution $ \pi (x) $.
Unlike a rejection sampling, MCMC does a random walk in a parametric space and concentrates on an important areas, which usually brings better efficiency.
Hamiltonian Monte Carlo(HMC)~\cite{N11} is introduced to provide better efficiency in converging to a desired distribution.
It adopts physical system dynamics to propose future states in the Markov chain.
The concept origins from Hamiltonian dynamics, which models the movement of an object.
Given the position $ x $ and the momentum $ q $ of an object, the system energy can be written as $ H(x,q) = U(x) + K(q) $, in which $ U(x) $ is the potential energy and $ K(q) $ is the kinetic energy.
The movement of the object (how $ x $ and $ q $ change over time) is determined by
$ \frac{dx}{dt} = \frac{\partial H}{\partial q} $ and $ \frac{dq}{dt} = - \frac{\partial H}{\partial x} $.
By the association between energy and a canonical distribution, a joint distribution for the states $ x $ and $ q $ can be created from a Hamiltonian dynamics, which is 
$ P(x, q) = \frac{1}{Z} \exp ( -H(x,q) / T ) = \frac{1}{Z} \exp ( -U(x) / T) \exp (-K(q) / T) $.
If we define $ U(x) = - \log ( \pi(x) T ) $, $ P(x, q) $ can be rewritten as $ P(x, q) = \pi (x) \frac{1}{Z} \exp (-K(q) / T) $.
Because the state $ x $ and the the momentum  $ q $ are independent, we can sample $ [ x ,q ] $ from $ P(x, q) $ and discard the momentum $ q $ (as an auxiliary variable).
The collected set of $ x $ is equivalent to sampling from $ \pi(x) $.
We are going to integrate this idea into our sampling framework of efficiently generating samples under kinetic constraints.

@@

>>>>

Initially, 
when no path has yet to be found, 
the samples need to be drawn from the entire state space~$\calX$.
However, once a path $\gamma$ is produced,  algorithms that seek \emph{high-quality paths}, which is the focus of this paper, can limit their sampling domain to a subset of~$\calX$  containing states that may produce higher-quality paths than~$\gamma$.
Following Gammell et al.~\cite{GSB14}, we call this subset the \emph{informed subset} and denote it~$\Cinf$.

@@

Unfortunately, both rejection sampling and HRS may produce a large number of rejected samples especially in high-dimensional spaces.
This may cause the search algorithm to spend most of its time trying to produces samples in~$\Cinf$ rather than explore it.
In this paper, we suggest an alternative approach to produce samples in the informed set \Cinf for a wide range of settings.
Our main insight is 
\os{this is a very important sentence!!!}
Specifically, our approach consists of two stages:
in the first, a random sample $x \in \calX$ is retracted to the boundary of~$\Cinf$ by running a gradient-descent algorithm;
in the second stage, this retracted sample $\tilde{x}$ is used to seed a Monte Carlo sampling chain which allows us to  produce samples that (approximately) cover~$\Cinf$  uniformly.

Our approach requires that the system has a solution to the two-point boundary value problem (2pBVP)~\cite{L06, H02} and that a gradient can be defined over the cost function.
Indeed, we demonstrate the efficiency of our approach on a wide variety of systems and show that it enables reducing the planning time by several orders of magnitute when compared to algorithms using rejection sampling or HRS.

@@

\subsection{Planning}
\label{subsec:planning}
Initial sampling-based algorithms such as RRT~\cite{LK01} and PRM~\cite{KSLO96} did not take into account the \emph{quality} of a path, given some optimization criteria, and only guaranteed to asymptotically return \emph{a} solution, if one exists.
Karaman and Frazzoli~\cite{KF11}, presented variants of PRM and RRT, named PRM* and RRT*, respectively that were shown to produce paths who's cost converges asymptotically to the minimal-cost path.
This is done by recognizing the underlying connections between stochastic sampling-based motion planning and the theory of random geometric graphs (see also~\cite{SSH16}).
Additional algorithms followed, increasing the converges rate by various techniques such as lazy dynamic programming~\cite{GSB15, JSCP15}, relaxing optimality to near-optimality~\cite{DB14, SH16} and more.

Many of the algorithms mentioned require solving a two-point boundary value problem (2pBVP) to perform exact and optimal connections between vertices in the roadmap.
For holonomic robots, these are simply straight lines in the state space, but for kinodynamic sytems with arbitrary cost functions,  computing an optimal trajectory between two states is non-trivial in general.

Xie et al.~\cite{XBPA15} use a variant of sequential quadratic programming (SQP) to solve 2pBVP and integrate it with BIT*~\cite{GSB15}.
Webb and van den Berg~\cite{WB13} use a fixed-final-state-free-final-time controller to solve the 2pBVP  with respect to a cost function that allows for balancing between the duration of the trajectory and the expended control effort.
Perez et al.~\cite{PPKKL12} propose a variant of RRT* that automatically defines a distance metric and node extension method by locally linearizing
the domain dynamics and applying linear quadratic regulation (LQR).

%Finally, we note that there have been many papers presenting solutions to the 2pBVP for specific systems such as
%
%Jeong Hwan Jeon, Raghvendra V. Cowlagi, Steven C. Peters, Sertac Karaman, Emilio Frazzoli, Panagiotis Tsiotras, Karl Iagnemma: Optimal motion planning with the half-car dynamical model for autonomous high-speed driving. ACC 2013: 188-193
%Dubins cars optimizng for both time and energy (RSS submission 107 by Shammas, Elie)

%possible additional papers on Kinodynamic RRT*:~\cite{KF10, GPPK13}

\subsection{Sampling}
\label{subsec:sampling}
There is a rich body of literature on how to produce samples that increase the efficiency of the planner in terms of finding a solution or producing high-quality solutions.
Heuristic approaches include
sampling on the medial axis~\cite{WAS99a, WAS99b, LTA03, YDLTA14},
sampling near the boundary of the obstacles~\cite{ABDJV98, YTEA12},
resampling along a given trajectory~\cite{APD11, AS11}
and more~\cite{US03, SWT09}

Of specific interest to our work are approaches that produce samples in the informed set~\Cinf.
As mentioned in Sec.~\ref{sec:intro} Gammel et al.~\cite{GSB14} describe an approach to sample uniformly in~\Cinf for the specific case of where $\calX = \R^d$ and when optimizing for path length.
To the best of our knowledge, the only method to produce samples in non-Euclidean spaces (other than rejection sampling) is HRS by Kunz et al.~\cite{KTC16}.

\subsection{MCMC}
\label{subsec:mcmc}
Markov Chain Monte Carlo (MCMC)~\cite{ADDJ03} is a framework of sampling from a given probability distribution by constructing a Markov chain of random transition toward a desired distribution $ \pi (x) $.
A set of samples are generated to approximate the desired distribution $ \pi (x) $.
Unlike a rejection sampling, MCMC does a random walk in a parametric space and concentrates on an important areas, which usually brings better efficiency.
Hamiltonian Monte Carlo(HMC)~\cite{N11} is introduced to provide better efficiency in converging to a desired distribution.
It adopts physical system dynamics to propose future states in the Markov chain.
The concept origins from Hamiltonian dynamics, which models the movement of an object.
Given the position $ x $ and the momentum $ q $ of an object, the system energy can be written as $ H(x,q) = U(x) + K(q) $, in which $ U(x) $ is the potential energy and $ K(q) $ is the kinetic energy.
The movement of the object (how $ x $ and $ q $ change over time) is determined by
$ \frac{dx}{dt} = \frac{\partial H}{\partial q} $ and $ \frac{dq}{dt} = - \frac{\partial H}{\partial x} $.
By the association between energy and a canonical distribution, a joint distribution for the states $ x $ and $ q $ can be created from a Hamiltonian dynamics, which is 
$ P(x, q) = \frac{1}{Z} \exp ( -H(x,q) / T ) = \frac{1}{Z} \exp ( -U(x) / T) \exp (-K(q) / T) $.
If we define $ U(x) = - \log ( \pi(x) T ) $, $ P(x, q) $ can be rewritten as $ P(x, q) = \pi (x) \frac{1}{Z} \exp (-K(q) / T) $.
Because the state $ x $ and the the momentum  $ q $ are independent, we can sample $ [ x ,q ] $ from $ P(x, q) $ and discard the momentum $ q $ (as an auxiliary variable).
The collected set of $ x $ is equivalent to sampling from $ \pi(x) $.
We are going to integrate this idea into our sampling framework of efficiently generating samples under kinetic constraints.

\section{Motivating example---Minimal Time Double Integrator}
We start by considering a simple yet important motion-planning problem

The double integrator is a simple yet im-
portant example that nicely illustrates the phase space. Suppose that a second-
order differential equation is given as q  ̈ = u, in which q and u are chosen from R

@@

\subsection{Minimal Time Double Integrator (MTDI)}
In this paper we demonstrate our approach for the specific case of a double integrator system when the objective function is minimizing the trajectory execution time.

Minimal Time Double Integrator
